---
title: "01 Data Analysis Basics"
author: Isabelle Langrock - Comm 522 
output: 
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
urlcolor: blue
---

Written by: Isabelle Langrock 
Comm 522: Introduction to Research Methods (Lab)
*Last Updated: August 2021* 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
```

# Objectives & Materials 

*This assignment provides an introduction to data analysis in R using the tidyverse package*

## Lab Objectives: 

1. Gain familarity with the RStudio environment and the basics of R 
2. Understand Categorical and Continuous Data
3. Learn and practice basic data analysis techniques (filtering, grouping, summarizing) and visualization skills.

The lab includes several questions for you to answer as well as several pre-written code sections to replicate. If you are new to R and/or programming in general, I urge you to type out each line of code, rather than using copy and paste. For additional help, reach out on the #r channel on slack.

## Getting Ready

Download this file as markdown file (.Rmd) and open in Rstudio. It's going to look a little differently than in the browser! Markdown files let you incorporate both code and text. You can run and manipulate the code included in this document, which you will need to do in order to complete this assignment. Basically, markdown files combine the script and output windows into one file. 

## Data Overview 

This lab will use the the covdata and covmobility packages from Kieran Healy (https://kjhealy.github.io/covdata/). A lot of data sets are included in this one package. Visit the codebook to see summaries of them all. https://kjhealy.github.io/covdata/articles/codebook.html

## Additional Resources 

- Chapter 5 of R for Data Science (https://r4ds.had.co.nz/transform.html) covers much of the same material (and more!)
- "Cheatsheets" for topics covered here
    - Rmarkdown: https://raw.githubusercontent.com/rstudio/cheatsheets/master/rmarkdown.pdf
    - dplyr: https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-transformation.pdf
    - Rstudio: https://raw.githubusercontent.com/rstudio/cheatsheets/master/rstudio-ide.pdf


# The Basics 

## Using R Markdown
A markdown file lets you include both text and code chunks in one file. You can then "knit" them into other file types (like html). 

This is what a code section looks like: 
```{r}
# when you use the "#" it's a text comment -- ie not code that will run. This is useful when
# annotating your code. 

# now try running these lines of code:  
2+2 
x<- c(1:5)
plot(x+2)

# to run your code you can either use the "Run" button at the top of this section of RStudio or you can highlight and use Command + Enter on your keyboard. 
```

You can insert code chunks using the insert button at the top of the window in R Studio. 

See this (https://raw.githubusercontent.com/rstudio/cheatsheets/master/rmarkdown.pdf) rmarkdown cheatsheet for a handy guide to all the possibilities of markdown! 


## Importing Packages & Data 

For this tutorial (and most analysis you will do in r) we need the tidyverse package. Some of you may have already installed this so we just need to open it.

To start working in R you will need to install the tidyverse packages. You only have to do this once, once installed you can call the package using library() everytime you need it. 
 R. 

We also need to install the data we will use. 

```{r Packages & Data }
# if you have not installed tidyverse, make sure to do so: 
# install.packages("tidyverse")
library(tidyverse)

# we'll install the data set directly from github. 
#devtools::install_github("kjhealy/covdata@main")
library(covdata)

#remotes::install_github("kjhealy/covmobility")
library(covmobility)
```


## Reviewing the data

We'll examine the google_mobility dataset, which notes daily aggregate changes in visits to different types of locations -- letting us examine how COVID impacts mobility. The pct_diff variable indicates a increase or decrease from a baseline average caclucated from a week prior to the pandemic. 

First, let's get a sense of the data: 

*Note: The data has gotten very large -- it might take a few minutes to install and load onto your computer!* 

```{r looking at the data}
data(google_mobility)
head(google_mobility, n=10) # Take a look at the first 10 observations. What do you think we would do to look at only the top 5? 
str(google_mobility) # Find a description of each variable 
summary(google_mobility) # Another description of each variable 
```

We can also view the full thing by running view(google_mobility) - but it's very large, so it's often easier to run these summary "looks". 

Some notes about the data: 

 - The accuracy and understanding of the place categories (type) do vary by region, so we should be hesitant with any firm conclusions that compare countries or urban and rural regions. 
 - The pct_diff variable measures change from a baseline median value that was caclculated based on data from January 2020 (before much of the world started social distancing). 
 - All data is collected from accounts that opted into sharing their Google location history with Google. There are many ways this might effect the generalizability of the findings.  

**What are the continuous and which are the categorical variables?** 

To get a count of variables and observations, we could count the variables from the summary commands above, but there are commands that do that as well:

```{r variables and observations}
ncol(google_mobility) # variables - each variable is a column
nrow(google_mobility) # observations - each observation is one row 
```

What are these observations of? In other words, what is our *unit of analysis*?

How many different cities, or metro areas, are included in the data set? How many different categories of type are there?

```{r}
length(unique(google_mobility$metro_area))

#How would you alter that function to find the number of categories included in the type variable?

#How many dates does the data set cover?

length(unique(google_mobility$date))

```

# Filtering Data 

Often, you might be concerned only with a specific set of cases from your data set. For example, maybe we just want to look at US cities. We'll need to filter and select the data we want. Let's save it as a new data frame, us_mobility. 

```{r Filtering Data}
# first we need to identify the country region code for the US. I'm pretty sure that States will be included in the country region so let's try that first. We'll use the str_detect funtion to identify  the word in the country_region column. 

google_mobility %>%
  filter(str_detect(country_region, "States"))

# aha there we go! Let's filter so we get a data set of only the US observations. We could use either country_region or country_region_code to do so. 


us_mobility <- google_mobility %>% 
  filter(country_region_code=="US")
  

# Filter the data so the only country included is Thailand. 

# How would you filter the data to only get observations for Brazil and Grocery Store? 

# We can also use filter on continuous variables, subsetting the data by a value. Here we filter the data to see only those pct_diff fewer than -100. 

google_mobility %>%
  filter(pct_diff< (-99))

# Filter the us data to find anything that had more than a 80 pct_diff increase
us_mobility %>%
  filter(pct_diff > 80)
```

# Grouping and summarizing data 

Another important skill in data science is the ability to summarize across different groups or subsets of data. We can make subsets using filter, as we did above, and then running analysis on the new dataframe, but if we want to aggregate across different categories, it's best to use the group_by function. 

```{r Grouping Data}
# Lets find a list of all the metropolitian areas included in our data set and count the number of times each type occurs 

google_mobility %>% 
  group_by(metro_area)%>%
  count(type)
  

# Are all 50 states included in our us_mobility data fame ?

# Sub_region_1 is equivalent to state for the US region so first lets rename the variable for clarity.
# then we'll count each state
us_mobility <- us_mobility %>%
  rename(state="sub_region_1")

us_mobility %>%
  count(state)

# what does the n in the table mean? 

# Now lets use our US dataframe and look at the mean change from the baseline (pct_diff) in each type 

us_mobility %>%
  group_by(type) %>%
  summarize(mean=mean(pct_diff, na.rm=TRUE))

# What category had the greatest increase across the US? 
# What category had greatest decrease? 


# Now find the mean and standard deviation [sd()] for each type for each state and assign it to the new variable state_type. 

```

# Data visualization 

Often the best way to get to know your data is by visualizing it! We'll review some of the basics of the ggplot package here and look at the distributions of the pct_diff variable for each type and state. 

```{r}
# first let's just plot the distributions of the type categories for our us data with a histogram

ggplot(us_mobility, aes(x=pct_diff, fill=type))+
  geom_histogram(binwidth = 20)
  
# That's not a very helpful visualization though. Let's facet the graph and make 6 different graphs. We'll also only look at PA data. 

us_mobility %>%
  filter(state== "Pennsylvania") %>%
  ggplot(us_mobility, mapping= aes(x=pct_diff))+
  geom_histogram(binwidth=20)+
  facet_wrap(~type)

# Still not the most useful. We can also use blox plots to look at the distribution of our data for each type. 

ggplot(us_mobility, aes(x=type, y=pct_diff, fill=type))+
  geom_boxplot()

# Make a box plot for the PA data: 

```

The data records changes over time. Let's visualize this. 

```{r}
ggplot(us_mobility, aes(x=date, y=pct_diff, group=state))+
  geom_line()+
  geom_smooth(aes(color=type, group=1), se=FALSE)+
  facet_wrap(~type)


# You can change the look of the graph with various themes. Try +theme_minimal() and +theme_dark. 
# Add a title, caption, and change the axis labels with +labs(x= " " , y= " ", title= " ", subtitle=""). 
```

Let's now look at the average change over time for each state for each type

```{r, fig.height=10, fig.width=7}
# first we want to make a new data frame that has an average for each state (& DC)

us_mobility %>%
  filter(state != "NA") %>%
  group_by(state, type, date) %>%
  summarise(ave_diff= mean(pct_diff, na.rm=TRUE))%>%
  ggplot(mapping=aes(x=date, y=ave_diff, color=type))+
  geom_line()+
  facet_wrap(~state, ncol=5)+
  theme_minimal()
```

***

# On your own: 
Complete the following tasks: 
*Feel free to use google, stackoverflow, the #r slack channel, or other resources to figure it out.*
*include your answers and your code* 

- Filter and assign a new data frame to a country or state of your choice from the data set. 
- Identify which type of location saw the greatest pct_diff. [HINT: try max()] What date did it occur on?
- Identify the mean, median, variance, and sd of all type categories for your chosen country. Use arrange() to sort the table from highest to lowest. 
- Plot the variations in pct_diff by type and date for your country. Make sure you include a title and descriptive axis labels. 
- Based on the graphs, how does your country compare to the US? Are there big differences? Based purely on conjecture (ie no need to look anything up!) propose a hypothesis about a similarities and/or differences. What data would one need to collect in order to test this hypothesis?
 

EXTRA CREDIT:
Using the apple_mobility dataset, examine the increase/decrease of use of certain transportation_types (as looked up in the Apple map app) over time for geo_type == "United States" and plot them. Identify a city and do the same. The baseline score is equal to 100 for this data set. 

```{r}
summary(apple_mobility)

```








